---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

#### This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day. 

##### Through this research we are attempting to use the existing data to 

## Loading and preprocessing the data



```{r,echo=TRUE }
require(ggplot2) 
require(plyr)
require(graphics)

fileRead <- na.omit(read.csv('activity.csv', header = TRUE, na.strings = "NA"))
dayAggregate <- aggregate(fileRead$steps, list(date = fileRead$date), sum)
colnames(dayAggregate) <- c("dates", "steps")


```

## What is mean total number of steps taken per day?

```{r,echo=TRUE}
dayAggregate$dates <- as.Date(dayAggregate$dates, "%Y-%m-%d") # we first ensure the date format is propper
dayAggregate$steps <- as.numeric(dayAggregate$steps) # and we also ensure is can be treated as numbers
dates <- dayAggregate$dates # we then fix the columns
steps <- dayAggregate$steps
qplot(dates, steps, geom = "auto") #to finally be able to graph the total mean number of steps per day
summary(dayAggregate)
```

#### AS seen in the 


## What is the average daily activity pattern?
```{r,echo=TRUE}
stepAggregate <- aggregate(fileRead$steps, list(interval = fileRead$interval), FUN = "mean")
colnames(stepAggregate) <- c("interval", "averageSteps")
plot(stepAggregate$interval, stepAggregate$averageSteps, type="l", xlab = "5-minute intervals", ylab = "Average number of steps taken")
summary(stepAggregate)
#aggregate(stepAggregate$interval, list(averageSteps = stepAggregate$averageSteps, max))
maximumAverage <- arrange(stepAggregate, stepAggregate$averageSteps, decreasing = TRUE)
head(maximumAverage,1)

```



## Imputing missing values
```{r,echo=TRUE}
missingValues <- read.csv('activity.csv', header = TRUE, na.strings = "NA")
sum(is.na(missingValues))


```




## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}

```

